---
title: "US EHE/ECE GridEX Data Loading"
format: html
editor: visual
description: "How to load GridEx data?"
author: "Pedram Fard"
date: "1/15/2024"
categories:
  - GridEX
  - tutorial
---

# Quick tutorial on loading the EHE/ECE GridEX data

## Recommended Package Setup

```{r setup}
library(readr)
library(viridis)
library(tidyverse)
library(sf)
```

## Loading the base spatial data set including Counties and States boundaries

```{r}
states_file_path <- url('https://github.com/epedram/us_ehe_ece/raw/main/iv_ehe_ece_data_integration/data/us49_states_geo_tigris.rds')

states_geo <- readRDS(states_file_path)[[1]]


counties_file_path <- url('https://github.com/epedram/us_ehe_ece/raw/main/iv_ehe_ece_data_integration/data/us49_counties_geo_tigris.rds')

counties_geo <- readRDS(counties_file_path)[[1]]

plot(counties_geo[1])
```

## Loading the aggregate extreme events data set

```{r load the aggreagte events summaries at State level}

events_catalog_file_path <- url('https://github.com/epedram/us_ehe_ece/raw/main/iv_ehe_ece_data_integration/data/States_events_by_type_sf_2008_2022.rds')

events_catalog_sf <- readRDS(events_catalog_file_path)[[1]] %>% 
  dplyr::filter(event_type == "Extreme Heat Event")

glimpse(events_catalog_sf)
```

## Example of merging States boundaries with summarized EHE data

```{r}
state_boundaris_ehe_catalog <- merge(states_geo,
          events_catalog_sf %>% st_drop_geometry(),
          by.x="GEOID",
          by.y="GEOID",
          all.x = TRUE,
          all.y = TRUE,
          suffix = c("","_sp")) %>% st_as_sf() 

glimpse(state_boundaris_ehe_catalog)
```

## Mapping the summarized EHE data merged with the States boundaries

```{r}
total_ehe_plot <- ggplot() +
    geom_sf(data = state_boundaris_ehe_catalog,
            aes(fill = total_event_days),
            lwd = .1) + 
    scale_fill_viridis(option="inferno", 
                       direction = -1,
                       na.value = NA) +

    labs(fill = "Total Number of EHE Days (2008-2022)") +

    theme_void() + 
    theme(legend.position = "bottom")

ggsave("./data/Total_Number_of_EHE_Days_2008_2022.jpg",
         plot = total_ehe_plot, dpi = 300,
         width = 22, height = 18, units = "cm")

total_ehe_plot
```


## For more information, check out:

-   [An Introduction to GridEX](intro_gridex.html)

-   [Detecting extreme heat events in California](extreme_heat_california.html)

-   [Github](https://github.com/epedram/us_ehe_ece/)
